---
title: "MPG trend research"
author: "Mariia Danilenko"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction  
We are interested in exploring the relationship between a set of variables and miles per gallon. We are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG?”  
"Quantify the MPG difference between automatic and manual transmissions."  

#Analysis  
## Using data
Our research is based on build in R dataset ***mtcars***.
We use **am** and **MPG** variables for the first version of calculations and all columns for the second.
```{r, message=FALSE}
library(datasets)
head(mtcars)
```

## First model
First linear regression model is based on only two variables **MPG** and **am**. As first step let's take a look at mean values aggregated by transmission type.
```{r}
boxplot(mtcars$mpg~factor(mtcars$am, labels=c("Automatic", "Manual")),
        col=c("lightblue", "darkblue"), xlab="Transmission type", 
        ylab="Mean MPG")

Means<-tapply(mtcars$mpg, factor(mtcars$am), mean)
names(Means)<-c("Automatic", "Manual")
Means
as.numeric(Means[2]-Means[1])
```
As we can see, Manual average **MPG** is 7.245 miles per galon higher than Automatic average **MPG**. 
We got the same result for the difference when constructed linear model on **AM** variable.      
```{r}
am_fit<-lm(mpg~am, data=mtcars)
summary(am_fit)$coef
```


## Advance model  
The best way to compare **MPG** for Manual and Automatic transmission types is to take a look at car models with same other technical features (CYL, WT and e.c.). However, we don't have this information and will try to explore linear regression model for **MPG** depends on all known variables.  
Also, we use function STEPAIC to obtain columns with the best result for different linear models.
```{r, message=FALSE}
library(MASS)
all_fit<-lm(mpg~., data=mtcars)
summary(all_fit)$coef
best_fit <- stepAIC(lm(mpg~., data=mtcars), trace=0)
summary(best_fit)$coef
```
Comparing three models:
```{r}
anova(am_fit, best_fit, all_fit)
```
S statistics says that adding **qsec** and **wt** is necessary, so we select **best_fit**. 
If we think about interpretation of variables, choosing *Weight* and *Transmission type* looks logically, but **QSec** is not. However, **QSec** has to depend on **Cyl** and **HP**, hence **MPG** depends on **Cyl** and **HP** too.       

## Residuals for selected model  
Let's plot residualts for the best_fit model:
```{r}
par(mfcol=c(2,2))
plot(best_fit)
```
There are no patterns at first, second and last plots.
Q-Q plot looks normal.
Based on these results we approbate **best_fit** model.   

#Conclusion  
**MPG** definetely depends on **am**. The rate differs according to a constructed model, but it is always will be positive: when you switch Automatic to Manual MPG grows. 
For the **best_fit**, where predictors are: **wt**, **am**, **qsec**, this rate is 2.9 mpg. 
